---
title: "Selecting and visualizing colour palettes"
author: "Monica" https://github.com/pr0nica/pr0nica
date: '2022-08-26'
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install.packages("ggplot2")
# install.packages("shiny")
# install.packages("RColorbrewer")
# install.packages("scales")
# install.packages("MetBrewer")
# install.packages("colorBlindness")


library(ggplot2)
library(shiny)

```

## RColorbrewer 
This palette has a collection of nice palettes with many of them being colorblind friendly.
The code below imports and shows how to display these colours, make a palette and visualise that palette in a plot. 

```{r  RcolorBrewer}
library(RColorBrewer) #import your package
display.brewer.all() #Displays all available palettes as plot
display.brewer.all(colorblindFriendly = T) # Show only colorblind friendly ones
display.brewer.pal(10, "RdYlBu") #Pick a palette and number of colours
```

There is also an interactive online tool for selecting colours similar to RColorbrewer:

https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3 

# Want to visualise the colour palettes (and hex code) before plotting? Here are some tricks:

### 1. "show_col"
This one gives you the colour and hex code together, if you want to compile your own palettes.
```{r show_col}
library(scales)

pal <- brewer.pal("RdYlBu", n=5) #define a palette
show_col(pal)

```


### 2. volcano image
I find this one useful to for gradients
```{r volcano}
pal <- brewer.pal("RdYlBu", n=5) #define a palette
image(volcano, col = pal) #Use volcano plot for a quick visualisation of the colours

```

### 3. ggplot
A cleaner way to display a palette with ggplot
```{r qplot}
pal <- brewer.pal("RdYlBu", n=5) #define a palette

qplot(x=1:5, y = 1, fill=factor(1:5), geom="tile") +
  scale_fill_manual(values = pal) +
  theme_void()+
  theme(legend.position="none") 



```



# Met Brewer palettes 

These palettes are inspired by famous artwork display in the New York Metropolitan museum. 
Here you can see all the palettes: 
https://twitter.com/blakerobmills/status/1478873962142593026?lang=ar-x-fm 

And here you can see the artwork they are created from:
https://github.com/BlakeRMills/MetBrewer/tree/main/PaletteImages 

```{r MetBrewer}
library(MetBrewer)
names(MetPalettes) #list all palettes
MetBrewer::colorblind.friendly("Thomas") #check if colorblind friendly
colorblind_palettes	 #list of all colorblind friendly palettes

```


#Change your palette directly in a figure!

Here's a simple tool to change colours for your palette one by one in shiny. I'm using colours from the MetBrewer as an example

```{r interactive, echo=TRUE, message=FALSE, warning=FALSE}
library(colorBlindness)

Hiroshige<- MetBrewer::met.brewer(name="Hiroshige",n=5,type="discrete")
show_col(Hiroshige)

#Make your input panel for colours (three variables, three colour options):
inputPanel(
  selectInput("palette1", label = "Setosa",
              choices = c(Hiroshige)), 
  
  selectInput("palette2", label = "Versicolor",
              choices = c(Hiroshige)),
  
  selectInput("palette3", label = "Virginica",
              choices = c(Hiroshige)),
  
  sliderInput("alpha", label = "transparency",
              min = 0.01, max = 1, value = 0.5, step = 0.05)
)


#Make your plot:
renderPlot({
  ggplot(iris)+
    geom_boxplot(aes(y=Sepal.Length,x=Species, group=Species, fill=Species),  col="black",alpha=input$alpha)+
    scale_color_manual(values=c(input$palette1, input$palette2,input$palette3), aesthetics = c("fill"))+
    theme_bw()

})

```

#Check your plot for colourblind-friendlyness and black-white print

The package "colorBlindness" allows you check how your colours look like with the two most common forms of color blindness (deuteranopia and protanopia). Below is a simple example, check out their guide for more details:

https://cran.r-project.org/web/packages/colorBlindness/vignettes/colorBlindness.html 

```{r colorBlindness1}
library(colorBlindness)

#Example with Hiroshige (my personal favourite)
p1<- ggplot(data=iris, aes(x=Species, y=Sepal.Length, fill=Species)) +
  geom_violin() +
  scale_fill_manual(values=met.brewer("Hiroshige", 3))+
  theme_bw()

cvdPlot(p1)

```

The plot above didn't look too good in BW, and while the colours are distinguishable I would prefer them to be more contrasting. Let's try some other colours with an interactive plot:

```{r colorBlindness2}
Homer1 <- met.brewer(name="Homer1",n=5,type="discrete")
show_col(Homer1)

#Example with Hiroshige (my personal favourite)

#Make your input panel for colours (three variables, three colour options):
inputPanel(
  selectInput("col1", label = "Setosa",
              choices = c(Homer1)), 
  
  selectInput("col2", label = "Versicolor",
              choices = c(Homer1)),
  
  selectInput("col3", label = "Virginica",
              choices = c(Homer1)),
  
  sliderInput("alpha", label = "transparency",
              min = 0.01, max = 1, value = 0.5, step = 0.05)
)


#Make your plot:
renderPlot({
  p1<- ggplot(iris)+
    geom_boxplot(aes(y=Sepal.Length,x=Species, group=Species, fill=Species),  col="black",alpha=input$alpha)+
    scale_color_manual(values=c(input$col1, input$col2,input$col3), aesthetics = c("fill"))+
    theme_bw()
cvdPlot(p1)

})




```


